---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# raoBust: Robust Rao (score) tests for Generalized Linear Models

<!-- badges: start -->
[![R-CMD-check](https://github.com/statdivlab/raoBust/workflows/R-CMD-check/badge.svg)](https://github.com/statdivlab/raoBust/actions)
[![codecov](https://codecov.io/github/statdivlab/raoBust/coverage.svg?branch=main)](https://app.codecov.io/github/statdivlab/raoBust)
<!-- badges: end -->

`raoBust`, at its core, gives you all the important information from `glm()`, but also with model misspecification-robust Rao tests (also called score tests) and Wald tests. 

Robust score tests have outstanding error rate performance in small samples, **and** when your data is not drawn from a parametric family (i.e., always). It's shocking to me (Amy) how well they perform. They have a reputation for being conservative in small samples, but I would argue that this is a *very good thing*. 

For now (because it's what I need for my work) I only have implementations for Poisson GLMs with log links. If you have another specific case you'd like to request, please let me know at [Issues](https://github.com/statdivlab/raoBust/issues) and label it as a "feature request". 

## Installation

You can install the development version of raoBust from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("statdivlab/raoBust")
```

## Example

This is a really silly example to show you the syntax. It looks just like `glm()` on the input side, but outputs a summary table that includes robust Wald and Rao tests as well as others. 

```{r example}
library(raoBust)
glm_test(dist ~ speed, data = cars, family=poisson(link="log"))
```

## Documentation 

We additionally have a `pkgdown` [website](https://statdivlab.github.io/raoBust/) that contains pre-built versions of our function [documentation](https://statdivlab.github.io/raoBust/reference/index.html) and an introductory [vignette](https://statdivlab.github.io/radEmu/articles/intro_raoBust.html).

## People

- Creator, maintainer: Amy D Willis
- Author: Sarah Teichman
- Author: David S Clausen
- Author: Shirley Mathur

All errors are Amy’s fault. 

## Naming

Score tests were pioneered by C.R. Rao, an outstanding 
statistician and methodologist. Score tests are sometimes, 
but increasingly rarely, called Rao tests. While Rao's work 
focused on non-robust score tests (not robust score tests) 
it is in tribute to Rao that this package is named. Many 
thanks to David Clausen for proposing and sharing the clever 
portmanteau `raoBust' for a package implementing robust score tests. 

**Comment from Amy** People of Color are consistently undervalued 
in science and mathematics, including in statistical methodology. 
The foundations of statistical methodology were built on 
the desire to quantitatively show the inferiority of People of Color 
and Jews. I believe that white supremacy continues to manifest 
in what names we give methods: we are more likely to name a method 
after a person if they are White, and more likely to name a method 
after its purpose, details or another acronym if the developer was 
a Person of Color. For this reason, I try to call score tests 
"Rao tests". I invite you to join me. 